---
title: "Exercises"
format:
  html:
    code-fold: true
    code-tools: true
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
    number_sections: true
author:
  - name: Tobias Rüttenauer
    orcid: 	0000-0001-5747-9735
    email: t.ruttenauer@ucl.ac.uk
    url: https://ruettenauer.github.io/
    affiliations:
      - name: UCL Social Research Institute
        address: University College London
        city: London WC1H 0AL
---

### Required packages

```{r, message = FALSE, warning = FALSE, results = 'hide'}
pkgs <- c("plm", "feisr", "sandwich", "texreg", "tidyr", "haven", "dplyr", "ggplot2", "ggforce") 
lapply(pkgs, require, character.only = TRUE)
```

### Session info

```{r}
sessionInfo()

```
  
### Load data

For the purpose of this exercise, we will use a real-world data set. Instead of constructing our own data, we use a shortcut and use data from the replication package of @Hospido.2012. The replication package can be found [here](http://qed.econ.queensu.ca/jae/datasets/hospido001/).

This is an unbalanced panel with 32,066 observations and 2066 individuals for the period 1968–1993 of the PSID. It consists of male household heads aged 25–55 with at least 9 years of usable wages data.


```{r}
# Load stata file
data.df <- read_dta("_data/h-data.dta")

# Lets order this
names <- names(data.df)
names <- c("pid", "year", names[-which(names %in% c("pid", "year"))])
data.df <- data.df[, names]

data.df <- data.df[order(data.df$pid, data.df$year), ]

```


| variable name   |   description|
|-----------------|-------------------------------------------------------|
| pid             |   INDIVIDUAL IDENTIFIER|
| year            |   YEAR OF INTERVIEW|
| age             |   AGE OF INDIVIDUAL|
| white           |   WHITE DUMMY|
| dropout         |   DROPOUT DUMMY|
| grad            |   GRADUATE DUMMY|
| college         |   COLLEGE DUMMY|
| married         |   MARRIED DUMMY|
| child           |   NUMBER OF CHILDREN|
| fsize           |   FAMILY SIZE|
| hours           |   YEARLY HOURS OF WORK|
| logwages        |   LOG OF REAL ANNUAL WAGES|
| changejob       |   JOB CHANGE DUMMY|
| ten1            |   TENURE DUMMY less than a year|
| ten2            |   TENURE DUMMY a year|
| ten3            |   TENURE DUMMY 2-3 years|
| ten4            |   TENURE DUMMY 4 through 9 years|
| ten5            |   TENURE DUMMY 10 through 19 years|
| ten6            |   TENURE DUMMY 20 years or more|
| profes          |   PROFESSIONAL, TECHNICAL, AND KINDRED WORKERS DUMMY|
| admin           |   MANAGERS AND ADMINISTRATORS DUMMY|
| sales           |   CLERICAL AND SALES WORKERS DUMMY|
| crafts          |   CRAFTSMAN AND KINDRED WORKERS DUMMY|
| operat          |   OPERATIVES WORKERS DUMMY|
| servic          |   LABORERS AND SERVICES WORKERS DUMMY|
| smsa            |   SMSA (Standard Metropolitan Statistical Area) DUMMY|
| neast           |   NORTH-EAST DUMMY|
| ncentr          |   NORTH-CENTRAL DUMMY|
| south           |   SOUTH DUMMY|
| west            |   WEST DUMMY|




# Exercise 1

Download and load the data. 

Have a look at the data. 

 * How many observations do we have in 1968? How many in 1990?


 * What is the average age in 1968? What was it in 1984? And how is this possible?


 * At which age did individual with ID "5790002" become father?


# Exercise 2

Just to play a little bit around with the data, let us estimate some models.

* What is the correlation between age and wage? Please use different estimators to determine different types of correlations.


# Exercise 3

Can we use this dataset to replicate our earlier analysis on the marital wage premium? What might be a problem here? (tip: have a look a the marriage variable).

However, we do something similar: We want to investigate if there is a fatherhood wage premium? In other words, do men experience an increase in wages when they become fathers?

    * Restrict the age at the start (first wage) to people aged 25-35

    

    * Use number of children to construct a binary indicator of wether there is a child in the household or not
    

    
    * Make sure we start only with men who are not yet fathers in the first period.
    


__Note: this feels like dropping a lot of information! However, it makes sense if we want to correctly identify the effect of interest.__

    * Do we need to drop observations where people go from child to no child?
    


    * Calculate the effect of having a child on the wage of men (including controls if reasonable). 
    
    
    
    * Calculate effects for POLS, RE, and FE (if you have some extra time, also FEIS). (#Hint: feis needs a class `data.frame` as input data)
    
    
    
    * Compare using cluster robust standard errors (and screenreg).


  
    * Interpret the results 


  
   * Try to perform a placebo test: what happens if you use the "lead" of becoming father. Why is this an interesting test?
   

# Exercise 4

Can we use one of the new event-study approaches, such as the Callaway and SantAnna estimator?

    * Preprocess data (treatment group and timing indicator)
    
    * Estimate the model using `att_gt`
    
    * Show the group-time specific estimates (use `method=ipw` and only restricted set of controls)
    
    * Interpret the results 

# References
  