---
title: "Exercise"
author: "Tobias Rüttenauer"
date: "December 03, 2021"
output_dir: docs
output: 
  html_document:
    theme: flatly
    highlight: haddock
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 2
theme: united
bibliography: Panel.bib
link-citations: yes
---

### Required packages

```{r, message = FALSE, warning = FALSE, results = 'hide'}
pkgs <- c("plm", "feisr", "panelView", "texreg", "tidyr", "haven", "dplyr", "ggplot2", "ggforce") 
lapply(pkgs, require, character.only = TRUE)
```

### Session info

```{r}
sessionInfo()

```
  
### Load data

For the purpose of this exercise, we will use a real-world data set. However, instead of constructing our own data, we use a shortcut and use data from the replication package of @Hospido.2012. The replication package can be found [here](http://qed.econ.queensu.ca/jae/datasets/hospido001/).

THis is an unbalanced panel with 32,066 observations and 2066 individuals for the period 1968–1993 of the PSID. It consists of male heads aged 25–55 with at least 9 years of usable wages data.


```{r}
# Load stata file
data.df <- read_dta("_data/h-data.dta")

# Lets order this
names <- names(data.df)
names <- c("pid", "year", names[-which(names %in% c("pid", "year"))])
data.df <- data.df[, names]

data.df <- data.df[order(data.df$pid, data.df$year), ]

```


| variable name   |   description|
|-----------------|-------------------------------------------------------|
| pid             |   INDIVIDUAL IDENTIFIER|
| year            |   YEAR OF INTERVIEW|
| age             |   AGE OF INDIVIDUAL|
| white           |   WHITE DUMMY|
| dropout         |   DROPOUT DUMMY|
| grad            |   GRADUATE DUMMY|
| college         |   COLLEGE DUMMY|
| married         |   MARRIED DUMMY|
| child           |   NUMBER OF CHILDREN|
| fsize           |   FAMILY SIZE|
| hours           |   YEARLY HOURS OF WORK|
| logwages        |   LOG OF REAL ANNUAL WAGES|
| changejob       |   JOB CHANGE DUMMY|
| ten1            |   TENURE DUMMY less than a year|
| ten2            |   TENURE DUMMY a year|
| ten3            |   TENURE DUMMY 2-3 years|
| ten4            |   TENURE DUMMY 4 through 9 years|
| ten5            |   TENURE DUMMY 10 through 19 years|
| ten6            |   TENURE DUMMY 20 years or more|
| profes          |   PROFESSIONAL, TECHNICAL, AND KINDRED WORKERS DUMMY|
| admin           |   MANAGERS AND ADMINISTRATORS DUMMY|
| sales           |   CLERICAL AND SALES WORKERS DUMMY|
| crafts          |   CRAFTSMAN AND KINDRED WORKERS DUMMY|
| operat          |   OPERATIVES WORKERS DUMMY|
| servic          |   LABORERS AND SERVICES WORKERS DUMMY|
| smsa            |   SMSA (Standard Metropolitan Statistical Area) DUMMY|
| neast           |   NORTH-EAST DUMMY|
| ncentr          |   NORTH-CENTRAL DUMMY|
| south           |   SOUTH DUMMY|
| west            |   WEST DUMMY|



# Exercise 1

Can we use this dataset to replicate our earlier analysis on the marital wage premium? What might be a problem here?

```{r}

```


# Exercise 2

However, we do something similar: is there a fatherhood wage premium? In other words, do men experience an increase in wages when they become fathers?

    * Restrict the age at the start (first wage) to people aged 25-35
    
```{r}

```
    

    * Use number of children to contruct a binary indicator of wether the is a child in the household or not
    
```{r}

```
    
    * Make sure we start only with men who are not yet fathers.
    
```{r}

```

__Note: this feels like dropping a lot of information! However, it makes sense if we want to correctly identify the effect of interest.__

    * Do we need to drop observations where people go from child to no child?
    
```{r}

```    

    * Calculate the effect of having a child on th wage of men (including controls if reasonable). 
    
    * Calculate effects for POLS, RE, and FE (if you have some extra time, also FEIS). (#Hint for feis: it needs a class `data.frame` as input data)
    
```{r}

```
    
    
    * Compare using Texreg.

```{r}

```
  
    * Interpret the results 



  
   * Try to perform a placebo test: what happens if you use the "lead" of becoming father. Why is this an interesting test?
   
```{r}

```



# References
  